### 2019-01-02 writing/thesis/chapters/methods.tex
Add second optional argument to basis macros.
Advance theory chapter.

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][i]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\newcommand{\sbasis}[1][]{\Phi_{#1}}
\newcommand{\stbasis}[1][]{\Theta_{#1}}
\newcommand{\sbasisRef}[1][]{\hat{\sbasis[#1]}}
\newcommand{\stbasisRef}[1][]{\hat{\stbasis[#1]}}
\newcommand{\stestfunction}[1]{\sbasis[#1]}
\newcommand{\sttestfunction}[1]{\stbasis[#1]}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}

@@

\chapter{An ADER-DG scheme for the Navier-Stokes Equations}\label{chap:methods}
This chapter describes the numerical and physical background that is needed to simulate reacting fluids with good accuracy.
We start with a description of the arbitrary high order derivatives (\ader) discontinuous Galerkin (\dg) method with a focus on equations that contain both advective and diffusive terms.
We then seque into a short description of the \muscl\ scheme, a second order finite volume scheme.
This discussion of the discretization enables us to talk about the equation set in sufficient detail.

@@

We will later need basis functions over a space-cell 
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasisRef(\hat{\bm{x}}) &= \sum_i^{(N+1)^D - 1} \lagrange[i] (x_i)\\
  \sbasis(\bm{x}) &= \sbasisRef \left( \mapping(\bm{x}) \right)
  \end{split}
\end{align}
where $\sbasisRef$ denotes the basis function on the reference cell.
Similar we define a basis for a space-time element defined by $(0, 1)^D \times (t^n, t^{n+1})$
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef(\hat{\bm{x}}, \hat{t}) &= \sum_j^{N} \sum_i^{(N+1)^D - 1} \lagrange[j](x_j) \lagrange[i] (x_i)\\
  \stbasis (\bm{x}, t) &= \stbasisRef \left( \mapping(\bm{x}), \timeMapping(t) \right)
  \end{split}
\end{align}

\sidetitle{Predictor}
To derive the predictor, we take an approximation of our solution in a nodal basis that is defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis{l}(\bm{x}, t).
\end{equation}
Furthermore we assume that all quantities are given in this represantation.
We now multiply the conservation law by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \sttestfunction{k} \source(\stpredictor)
}}.
\end{equation}
We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
Note that this neglects the interaction with neighbouring cells, we account for this in the corrector step.
\todo{Replace $\bm{x}$ with a macro some other letter. Is also coordinate!}
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
-
\intdcell{
  \sttestfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
} = \\
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
&+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be by a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Collect into matrices and show scheme?}
To show the computational efficiency of our scheme, we follow the approach of~\cite{dumbser2008unified} and collect our integrals into matrices\todo{Source is Dominics \aderdg-document.}.

\newcommand{\pleft}{\bm{L}}
\newcommand{\prightsol}{\bm{r}}
\newcommand{\prightpred}{\bm{w}}
\begin{align}
  \label{eq:predictor-matrices}
  \pleft &= \ldots \\
  \prightsol &= \ldots \\
  \prightpred &= \ldots
\end{align}
Using these matrices, we can solve for the space-time-predictor with the iterative scheme
\begin{align}
  \label{eq:predictor-iterative}
  \pleft \stpredictor &= \prightsol (\dsol) - \prightpred(\stpredictor) \\
 \stpredictor &= \pleft^{-1} \left( \prightsol (\dsol) - \prightpred(\stpredictor) \right).
\end{align}
For details and proof of convergence for the linear case, see~\cite{dumbser2008unified}.

We then prepare the space-time-predictor for use in the corrector step.
Due to the linearity of the Riemann solver, we can exchange time and space integration and thus only need to consider the Riemann problem for time-averaged quantities.
We can compute this for the space-time-predictor by
\begin{align}
  \ldots = \ldots,
\end{align}
the procedure is similar for all other quantities (Q, gradQ, fluxes).

Finally, we extrapolate all quantities to the boundary.
This is needed for the Riemann solver and for the reconstruction of boundary values.
We achieve this by
\begin{align}
  \label{eq:boundary-extrapolation}
  \ldots = \ldots.
\end{align}

\sidetitle{Boundary extrapolation and stuff}
\todo{Explain that stpredictor and fluxes are saved and extrapolated to boundary.
This is important because we extrapolate the gradient as well!}
Extrapolate unknowns to boundary, convert from our basis to basis on faces and so on

\sidetitle{Corrector}
First, we multiply the system \cref{eq:conservation-law-gradient} by a test function $\stestfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
Note that the testfunctions
\begin{equation}
  \label{eq:cell-approx-space}
  q(\bm{x})_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}).
\end{equation}

here do not depend on time but only on the spatial dimensions.
We arrive yet again at the weak formulation of the \pde\
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction{k} \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction{k} S(\Q, \bm{x}, t)
}}.
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with the spacetime-predictor $\stpredictor (\bm{x},t)$ and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction{k} \sbasis[l]
}}
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcell{
      \stestfunction{k} \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\stestfunction{k}} \cdot  \flux(\stpredictor, \gradient{\stpredictor}) % todo correct gradient?
}}\right)
&=
\left(\intdt{\intdcell{
      \stestfunction{k} \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term is our mass matrix with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
\begin{equation}
  \label{eq:mass-matrix}
  \massMatrix[kl] = \massMatrixDef = \volume \underbrace{\quadWeight[k] \delta_{kl}}_{\text{diagonal}}
\end{equation}
which is diagonal thanks to the orthogonal basis \pcref{eq:basis-orthogonality} and thus trivial to invert.
\newcommand{\sumbasis}{\sum_i^{(N + 1)^d - 1}}
We can then insert the definition of our basis function into our equation and arrive at
\begin{align}
  \massMatrix \left( \dsol^{t+1} - \dsol^{t} \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
\begin{split}
  a &= \sumbasis F_n \intdcell{
\sbasis[n] \gradient{\sbasis[n]}
  } \\
  b &= \ldots \\
  s_n &= \sumbasis \intdcell{
s \sbasis[n]
  }.
\end{split}
\end{align}
We can solve this for the new unknows $\bm{u}^{n+1}$
\begin{align}\label{eq:update-predictor}
\begin{split}
  \bm{u}^{n+1} &= \bm{u}^{n} + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix^{-1} \left( \bm{a} - \bm{b} + \bm{s} \right)
\end{split}
\end{align}
which is clearly of the form of a one-step scheme.

@@

We then use the slope to reconstruct the value of the boundary.
In the following $\cellAvg^{\pm x}$ is the average of the left ($+$) or right ($-$) cell boundary.
Similar, $\cellAvg^{\pm y}$ is the value at the top/bottom cell boundary.
These so called boundary extrapolated values are given by
\newcommand{\extrapolatedCellAvg}[3][i,j]{\cellAvg[#1]^{#3 #2} = \cellAvg #3 \frac{1}{2} \slope{#2}}
\begin{equation}
\begin{alignedat}{2}
& \extrapolatedCellAvg{x}{-} , \qquad && \extrapolatedCellAvg{x}{+}, \\
& \extrapolatedCellAvg{y}{-} , \qquad && \extrapolatedCellAvg{y}{+}.
\end{alignedat}
\end{equation}
We also use the slopes defined in \cref{eq:slopes} to estimate the gradient of $\cellAvg$ in each cell by the block matrix\todo{Notation!}
\begin{equation}
  \label{eq:muscl-gradient}
  \gradCellAvg = \left( \slope{x} \bigg\rvert \slope{y} \right),
\end{equation}
which is of course an abuse of notation insofar as it is not the gradient of the constant cell value but rather of the linear reconstruction \pcref{eq:slopes}.
\sidetitle{Time Evolution}
To achieve second order in time, we evolve the boundary extrapolated values in time by
\todo{Correct index for gradient of cell average}
\todo{Actually compare with implementation. Currently gradient is ignored here!}
\todo{Describe splliting of $\flux$ into $\fluxX, \fluxY$}
\newcommand{\evolvedCellAvg}[2][i,j]{\hat{U}_{#1}^{#2}}
\begin{equation}\label{eq:muscl-time-evolution}
  \begin{split}
  \forall_{k \in [-x, +x, -y, +y]}:  \hat{U}_{ij}^k = \cellAvg^k &+
  \frac{\Delta t}{2 \Delta x} \left[ \fluxX(\cellAvg^{-x}, \gradCellAvg) - \fluxX(\cellAvg^{+x}, \gradCellAvg) \right] \\&+
  \frac{\Delta t}{2 \Delta y} \left[ \fluxY(\cellAvg^{-y}, \gradCellAvg) - \fluxY(\cellAvg^{+y}, \gradCellAvg) \right].
  \end{split}
\end{equation}

Finally, the update can be described by\todo{Double check update}
\begin{align}
  \begin{split}
    \cellAvg (t^{n+1}) &= \cellAvg (t^n)
    \\ &+
  \frac{\Delta t}{\Delta x}
    \Riemann \left(
      \evolvedCellAvg[i-1,j]{+x}, \gradCellAvg[i-1,j]; \ %
      \evolvedCellAvg[i,j\phantom{-1}]{-x}, \gradCellAvg[i,j\phantom{-1}]
    \right)
    \\ &-
  \frac{\Delta t}{\Delta x}
    \Riemann \left(
      \evolvedCellAvg[i,j\phantom{+1}]{+x}, \gradCellAvg[i,j\phantom{+1}]; \ %
      \evolvedCellAvg[i+1,j]{-x}, \gradCellAvg[i+1,j]
    \right)
    \\ &+
  \frac{\Delta t}{\Delta y}
    \Riemann \left(
      \evolvedCellAvg[i,j-1]{+y}, \gradCellAvg[i,j-1]; \ %
      \evolvedCellAvg[i,j\phantom{-1}]{-y}, \gradCellAvg[i,j\phantom{-1}]
    \right)
    \\ &-
  \frac{\Delta t}{\Delta y}
    \Riemann \left(
      \evolvedCellAvg[i,j\phantom{+1}]{+y}, \gradCellAvg[i,j\phantom{+1}]; \ %
      \evolvedCellAvg[i,j+1]{-y}, \gradCellAvg[i,j+1]
    \right)    
  \end{split}
\end{align}

@@

\todo{Gradient on reference cell, Gaussian quad., inline}
\todo{Definition for FV schemes (maybe just mapping to DG-Cells)}
\newcommand{\gobs}{\operatorname{G}}
\newcommand{\mean}{\mu}
\newcommand{\std}{\sigma}
\newcommand{\variance}{\std^2}
\newcommand{\gobsCount}{n}
A simple way of determining outliers is by how many standard deviations it differs from the mean value.
We thus want to compute the mean and variance over all grid cells.
To minimize intranodal communications and to simplify the implementation, we do this by performing only pairwise merge operations.
The simplest way to compute this would be by using the textbook definitions of mean $\mean$ and variance $\variance$ of a random variable $X$
\newcommand{\expectation}{\mathbb{E}}
\begin{align}
  \begin{split}
    \mean [X] &= \expectation \left[ X \right]\\
    \variance [X] &= \expectation \left[ X^2 \right] - \expectation \left[ X \right]^2.
  \end{split}
\end{align}
The naive algorithm is numerically unstable when the variance is orders of magnitudes smaller than the mean as this can lead to catastrophic cancellations.
We thus compute the mean and variance with the parallel algorithm of~\cite{chan1982updating}.
To do this, we need to store three variables per computation unit:
the mean, the variance and the number of processed elements.
We collect these items in the vector $\gobs = (\mean, \variance, \gobsCount)$
We can then merge two pairs of observed variables with \cref{alg:merge-variance}.
This algorithm computes an estimate of the population variance using Bessel's correction, i.e.\ it returns the sample variance times $\nicefrac{n}{n-1}$.
In our case, the compute the variance over all grid cells.
Because of this, we multiply the estimate of \cref{alg:merge-variance} with $\nicefrac{n-1}{n}$ to undo the correction.

>>>>

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\NewDocumentCommand{\sbasis}{ O{} O{} }{\phi^{#1}_{#2}}
\NewDocumentCommand{\stbasis}{ O{} O{} }{ \Theta^{#1}_{#2}}
\NewDocumentCommand{\sbasisRef}{ O{} O{} }{\hat{\sbasis[#1][#2]}}
\NewDocumentCommand{\stbasisRef}{ O{} O{} }{\hat{\stbasis[#1][#2]}}
%\newcommand{\stestfunction}[1]{\sbasis[#1]}
\NewDocumentCommand{\stestfunction}{ O{} O{} }{\sbasis[#1][#2]}
\NewDocumentCommand{\sttestfunction}{ O{} O{} }{\stbasis[#1][#2]}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}

@@

\chapter{An ADER-DG scheme for the Navier-Stokes Equations}\label{chap:methods}
This chapter describes the numerical and physical background that is needed to simulate reacting fluids with good accuracy.
We start with a description 
We start with a description of the arbitrary high order derivatives (\ader) discontinuous Galerkin (\dg) method with a focus on equations that contain both advective and diffusive terms.
We then seque into a short description of the \muscl\ scheme, a second order finite volume scheme.

@@

We will later need basis functions over a space-cell 
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasisRef(\hat{\bm{x}}) &= \sum_i^{(N+1)^D - 1} \lagrange[i] (x_i)\\
  \sbasis(\bm{x}) &= \sbasisRef \left( \mapping(\bm{x}) \right)
  \end{split}
\end{align}
where $\sbasisRef$ denotes the basis function on the reference cell.
Similarly, we define a basis for a space-time element defined by $(0, 1)^D \times (t^n, t^{n+1})$
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef(\hat{\bm{x}}, \hat{t}) &= \sum_j^{N} \sum_i^{(N+1)^D - 1} \lagrange[j](x_j) \lagrange[i] (x_i)\\
  \stbasis (\bm{x}, t) &= \stbasisRef \left( \mapping(\bm{x}), \timeMapping(t) \right)
  \end{split}
\end{align}

\sidetitle{Predictor}
To derive the predictor, we take an approximation of our solution in a nodal basis that is defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis[][l](\bm{x}, t).
\end{equation}
Furthermore we assume that all quantities are given in this represantation.
We now multiply the conservation law by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \sttestfunction{k} \source(\stpredictor)
}}.
\end{equation}
We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
This corresponds to upwinding in time~\cite{dumbser2008unified}.
Note that this neglects the interaction with neighbouring cells.
We account for this in the corrector step.
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
= \\
\intdcell{
  \sttestfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
}
&+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be by a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Collect into matrices and show scheme?}
To show the computational efficiency of our scheme, we follow the approach of~\cite{dumbser2008unified} and collect our integrals into matrices\todo{Source is Dominics \aderdg-document.}.

\newcommand{\pleft}{\bm{L}}
\newcommand{\prightsol}{\bm{r}}
\newcommand{\prightpred}{\bm{w}}
\begin{align}
  \label{eq:predictor-matrices}
  \pleft^{\cell}_{(n v l), (n' v' l')} &=
\intdcell{
  \sttestfunction[\cell][v,l,n](1) \stbasis[\cell][v,l', n']
}
-
\intdt{\intdcell{
  \left( \pdv{}{t} \sttestfunction[\cell][v,l,n] (t) \right) \stbasis[\cell][v,l',n']
}}                             
\\
\prightsol^{\cell}_{v, l,n} &=
\intdcell{
  \sttestfunction[\cell][v,l,n] (0) \sbasis[\cell][v,n'] \dsol[v,n']^{\cell}
}
+
\intdt{\intdcell{
  \sttestfunction[\cell][v,l,n] \stbasis[\cell][v,l',n'](t) S_{v,l',n'}(\stpredictor[]^{\cell, i})
}}                             
  \\
\prightpred^{\cell}_{v,l,n} &= 
F^{k}_{v,l',n'}
\intdt{\intdcell{
  \left( \gradient{\sttestfunction[\cell][v,l,n]} \right) \stbasis[\cell][v,l',n'](t) S_{v,l',n'}(\stpredictor[]^{\cell, i})
}}
\end{align}
Using these matrices, we can solve for the space-time-predictor with the iterative scheme
\begin{align}
  \label{eq:predictor-iterative}
  \pleft \stpredictor[]^{\cell, i+1} &= \prightsol (\dsol) - \prightpred(\stpredictor[]^{\cell, i}) \\
 \stpredictor[]^{\cell, i+1} &= \pleft^{-1} \left( \prightsol (\dsol) - \prightpred(\stpredictor[]^{\cell, i}) \right).
\end{align}
For details and proof of convergence for the linear case, see~\cite{dumbser2008unified}.

\sidetitle{Preparing for corrector}
We then prepare the space-time-predictor for use in the corrector step.
Due to the linearity of the Riemann solver, we can exchange time and space integration and thus only need to consider the Riemann problem for time-averaged quantities.
% We can compute this for the space-time-predictor by
% \begin{align}
%   \ldots = \ldots,
% \end{align}
% the procedure is similar for all other quantities (Q, gradQ, fluxes).
We then extrapolate all quantities to the boundary.
This is needed for the Riemann solver and for the reconstruction of boundary values.
We accomplish this by evaluating the function at the degrees of freedom on the boundary.
Thanks to the interpolating property of the nodal basis~\pcref{eq:basis-interpolating}, this reduces to a change of basis which can be precomputed on the reference element.
We compute time-averaged and boundary extrapolated versions of the space-time-predictor, the degrees of freedom and their gradient, and the fluxes.

\sidetitle{Corrector}
The final step of the \aderdg{}-scheme is the corrector.
For this step, we write the approximation of a function $q$ in a cell as
\begin{equation}
  \label{eq:cell-approx-space}
  q(\bm{x})_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}),
\end{equation}
where the basis, unlike in \cref{eq:cell-approx-time}, does not depend on the time.
We then multiply the system \cref{eq:conservation-law-gradient} by a test function $\stestfunction{i}$ and integrate over the space-time volume $(\cell \times [t^n, t^{n+1}])$.
We arrive yet again at the weak formulation of the \pde\
\begin{equation}
  \label{eq:weak-pde}
\intdt{\intdcell{
\stestfunction{k} \pdv{\Q}{t}
}}
+
\intdt{\intdcell{
    \stestfunction{k} \left( \div{F(\Q, \gradQ } \right)
}}
=
\intdt{\intdcell{
    \stestfunction{k} S(\Q, \bm{x}, t)
}}.
\end{equation}
\todo{Include gradient everywhere!}
\todo{Double check the following description}
We now replace the solution $\bm{Q}$ with the spacetime-predictor $\stpredictor (\bm{x},t)$ and write it as a polynomial using the representation of \cref{eq:cell-approx}.
Integrate first by parts in time (note basis here not defined over time), flux divergence by parts in space.
\newcommand{\massMatrixDef}{\intdcell{
  \stestfunction[\cell][k] \sbasis[\cell][l]
}}
\begin{align}
\begin{split}
\label{eq:corrector}
\left(
\massMatrixDef
\right)
(\bm{u^{n+1} - u^{n}})
&+
\left(\intdt{\intdcellb{
      \stestfunction[k] \Riemann(\stpredictor^-, \stpredictor^+) \cdot \normal
}}\right)
-\\
\left(\intdt{\intdcell{
    \gradient{\stestfunction[k]} \cdot  \flux(\stpredictor, \gradient{\stpredictor}) % todo correct gradient?
}}\right)
&=
\left(\intdt{\intdcell{
      \stestfunction[k] \source(\stpredictor)
}}\right)
\end{split}
\end{align}
The first term is our mass matrix with entries
\newcommand{\massMatrix}[1][]{\bm{M}_{#1}}
\begin{equation}
  \label{eq:mass-matrix}
  \massMatrix[kl] = \massMatrixDef = \volume \underbrace{\quadWeight[k] \delta_{kl}}_{\text{diagonal}}
\end{equation}
which is, due to the orthogonal basis \pcref{eq:basis-orthogonality}, diagonal and thus trivial to invert.
\newcommand{\basisSize}{(N+1)^d - 1}
\newcommand{\sumbasis}{\sum_i^{\basisSize}}
We can then insert the definition of our basis function into our equation and arrive at
\begin{align}
  \massMatrix \left( \dsol^{t+1} - \dsol^{t} \right) &=
  \Delta t \left( \bm{a} - \bm{b} + \bm{s} \right) 
  \intertext{with}
\begin{split}
a^{\cell}_{v,n} &= F^{\cell}_{n'} \intdt{\intdcell{
  \sbasis[\cell][v,n'] \gradient{\sbasis[\cell][v,n]}(t)
}}
\\
b &= \ldots \\
s^{\cell}_n &=
\intdcell{
  S^{\cell}_v \sbasis[\cell][v,n]
}
,
\end{split}
\end{align}
where we evaluated the time-integrals by using the fact that our quantities are time-averaged.
We can solve this for the new unknows $\bm{u}^{n+1}$
\begin{align}\label{eq:update-predictor}
\begin{split}
  \bm{u}^{n+1} &= \bm{u}^{n} + \Delta t (\operatorname{update}) \\
  \operatorname{update} &= \massMatrix^{-1} \left( \bm{a} - \bm{b} + \bm{s} \right)
\end{split}
\end{align}
which is clearly of the form of a one-step scheme.

@@

We then use the slope to reconstruct the value of the boundary.
In the following $\cellAvg^{\pm x}$ is the average of the left ($+$) or right ($-$) cell boundary.
Similar, $\cellAvg^{\pm y}$ is the value at the top/bottom cell boundary.
These so called boundary extrapolated values are given by
\newcommand{\extrapolatedCellAvg}[3][i,j]{\cellAvg[#1]^{#3 #2} = \cellAvg #3 \frac{1}{2} \slope{#2}}
\begin{equation}
\begin{alignedat}{2}
& \extrapolatedCellAvg{x}{-} , \qquad && \extrapolatedCellAvg{x}{+}, \\
& \extrapolatedCellAvg{y}{-} , \qquad && \extrapolatedCellAvg{y}{+}.
\end{alignedat}
\end{equation}
We also use the slopes defined in \cref{eq:slopes} to estimate the gradient of $\cellAvg$ in each cell by the block matrix\todo{Notation!}
\begin{equation}
  \label{eq:muscl-gradient}
  \gradCellAvg = \left( \slope{x} \bigg\rvert \slope{y} \right),
\end{equation}
which is of course an abuse of notation insofar as it is not the gradient of the constant cell value but rather of the linear reconstruction \pcref{eq:slopes}.
\sidetitle{Time Evolution}
To achieve second order in time, we evolve the boundary extrapolated values in time by
\todo{Correct index for gradient of cell average}
\todo{Actually compare with implementation. Currently gradient is ignored here!}
\todo{Describe splliting of $\flux$ into $\fluxX, \fluxY$}
\newcommand{\evolvedCellAvg}[2][i,j]{\hat{U}_{#1}^{#2}}
\begin{equation}\label{eq:muscl-time-evolution}
  \begin{split}
  \forall_{k \in [-x, +x, -y, +y]}:  \hat{U}_{ij}^k = \cellAvg^k &+
  \frac{\Delta t}{2 \Delta x} \left[ \fluxX(\cellAvg^{-x}, \gradCellAvg) - \fluxX(\cellAvg^{+x}, \gradCellAvg) \right] \\&+
  \frac{\Delta t}{2 \Delta y} \left[ \fluxY(\cellAvg^{-y}, \gradCellAvg) - \fluxY(\cellAvg^{+y}, \gradCellAvg) \right].
  \end{split}
\end{equation}
Note that we do not need to consider neighboring cells for this step.

Finally, the update can be described by\todo{Double check update}
\begin{align}
  \begin{split}
    \cellAvg (t^{n+1}) &= \cellAvg (t^n)
    \\ &+
  \frac{\Delta t}{\Delta x}
    \Riemann \left(
      \evolvedCellAvg[i-1,j]{+x}, \gradCellAvg[i-1,j]; \ %
      \evolvedCellAvg[i,j\phantom{-1}]{-x}, \gradCellAvg[i,j\phantom{-1}]
    \right)
    \\ &-
  \frac{\Delta t}{\Delta x}
    \Riemann \left(
      \evolvedCellAvg[i,j\phantom{+1}]{+x}, \gradCellAvg[i,j\phantom{+1}]; \ %
      \evolvedCellAvg[i+1,j]{-x}, \gradCellAvg[i+1,j]
    \right)
    \\ &+
  \frac{\Delta t}{\Delta y}
    \Riemann \left(
      \evolvedCellAvg[i,j-1]{+y}, \gradCellAvg[i,j-1]; \ %
      \evolvedCellAvg[i,j\phantom{-1}]{-y}, \gradCellAvg[i,j\phantom{-1}]
    \right)
    \\ &-
  \frac{\Delta t}{\Delta y}
    \Riemann \left(
      \evolvedCellAvg[i,j\phantom{+1}]{+y}, \gradCellAvg[i,j\phantom{+1}]; \ %
      \evolvedCellAvg[i,j+1]{-y}, \gradCellAvg[i,j+1]
    \right).
  \end{split}
\end{align}

@@

\todo{Gradient on reference cell, Gaussian quad., inline}
\todo{Definition for FV schemes (maybe just mapping to DG-Cells)}
\newcommand{\gobs}{\operatorname{G}}
\newcommand{\mean}{\mu}
\newcommand{\std}{\sigma}
\newcommand{\variance}{\std^2}
\newcommand{\gobsCount}{n}
A simple way of determining outliers is by how many standard deviations it differs from the mean value.
We thus want to compute the mean and variance over all grid cells.
To minimize intranodal communications and to simplify the implementation, we do this by performing only pairwise merge operations.
The simplest way to compute this would be by using the textbook definitions of mean $\mean$ and variance $\variance$ of a random variable $X$
\newcommand{\expectation}{\mathbb{E}}
\begin{align}
  \begin{split}
    \mean [X] &= \expectation \left[ X \right]\\
    \variance [X] &= \expectation \left[ X^2 \right] - \expectation \left[ X \right]^2.
  \end{split}
\end{align}
The naive algorithm is numerically unstable when the variance is orders of magnitudes smaller than the mean as this can lead to catastrophic cancellations.
We thus compute the mean and variance with the parallel algorithm of~\cite{chan1982updating}.
To do this, we need to store three variables per computation unit:
the mean, the variance and the number of processed elements.
We collect these items in the vector $\gobs = (\mean, \variance, \gobsCount)$
We can then merge two pairs of observed variables with \cref{alg:merge-variance}.
This algorithm computes an estimate of the population variance using Bessel's correction, i.e.\ it returns the sample variance times $\nicefrac{n}{n-1}$.
In our case, the compute the variance over all grid cells.
\todo{Std is still slightly biased!}
Because of this, we multiply the estimate of \cref{alg:merge-variance} with $\nicefrac{n-1}{n}$ to undo the correction.\todo{Is this actually correct? Doesn't bessel's correction come from the dofs ``used for the mean?''. (Doesnt really matter anyway.)}

\begin{algorithm}[ht]
  \begin{algorithmic}
% TODO: Special cases 
\Function{reduce-variance}{$\gobs_0, \gobs_1$}
\If{$\gobsCount_0 = 0$}
  \State\Return{$\mean_1, \variance_1, \gobsCount_1$}
\EndIf\
\If{$\gobsCount_1 = 0$}
  \State\Return{$\mean_0, \variance_0, \gobsCount_0$}
\EndIf\
  \Let{$\Delta$}{$\mean_1 - \mean_0$}  
  \Let{$\gobsCount_\Sigma$}{$\gobsCount_0 + \gobsCount_1$}
  \Let{$m_a$}{$\variance_0 (\gobsCount_0 - 1)$}
  \Let{$m_b$}{$\variance_1 (\gobsCount_1 - 1)$}
  \Let{$m_\Sigma$}{$\nicefrac{m_a + m_b + (\Delta^2 \gobsCount_0 \gobsCount_1)}{\gobsCount_\Sigma}$}
  \Let{$\mean_\Sigma$}{$\mean_0  \gobsCount_0 + \mean_1 \gobsCount_1$}
  \State\Return{$
    \nicefrac{\mean_\Sigma}{\gobsCount_\Sigma},
    \nicefrac{m_\text{total}}{\gobsCount_\Sigma - 1},
    \gobsCount_\Sigma
    $}
\EndFunction\
  \end{algorithmic}
  \caption{\label{alg:merge-variance}
    Merging two sets of reduced mean and variance~\cite{chan1982updating}}
\end{algorithm}
Chebychev's inequality
\begin{equation}
  \label{eq:chebychev}
  \mathbb{P}(\vert X - \mu \vert \geq c \sigma) \leq \frac{1}{c^2},
\end{equation}
with probability $\mathbb{P}$,
holds for an arbitrary distribution with mean $\mu$ and standard deviation $\sigma$ probability $\mathbb{P}$~\cite{wasserman2004all}.
This motivates the following refinement criterion
\begin{equation}
  \label{eq:refinement-criterion}
  \operatorname{evaluate-refinement}(\Q, \mu, \sigma) =
  \begin{cases}
    \text{refine} & \text{if } \tv(\Q) \geq \mu + T_\text{refine} \sigma \\
    \text{delete} & \text{if } \tv(\Q) \leq \mu + T_\text{delete} \sigma \\
    \text{keep} & \text{otherwise}
    \end{cases}
\end{equation}
which refines when a cell has a significantly higher standard deviation.
In this equation, the constants $T_r$ and $T_d$ can be used to tailor the trade-off between the quality of approximation and computational cost.
They can be chosen freely, as long as $T_d < T_r$ holds.
Chebychev's inequality then guarantees that only a subset is marked for refinement.
Note that this inequality provides only a loose bound.
If one would be willing to assume a distribution of the indicator variable, tighter bounds can be derived.

\subsection{Limiting}\label{sec:limiting}
Higher order \textsc{dg} methods cannot cope with discontinuous solutions.
Even worse, in the case of non-linear fluxes, discontinuities can appear even from smooth initial data.
In addition, oscillations stemming from the high-order polynomical approximation can lead to wrong or unphyiscal data.
This is called Gibbs phenomenon.
A classical way of dealing with this problem is the usage of so called limiters.
The idea is to smooth out steep gradients and thus eliminate discontinuities~\cite{hesthaven2008nodal}.\todo{It doesnt work like this exactly.}
